---
title: "Maximum Mean Discrepancy"
subtitle: ""
author: "Frederick Boehm"
institute: ""
date: "March 11, 2021 (updated: `r Sys.Date()`)"
bibliography: mmd.bib
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, load_refs, include=FALSE, cache=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "alphabetic",
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
bib <- ReadBib("mmd.bib", check = FALSE)
```


## Maximum Mean Discrepancy: What is it? 

- Statistical method for inferring whether two samples have distinct underlying distributions, $P$ and $Q$ `r Citep(bib, "gretton2012kernel")`

- Idea: 
    - Find a smooth function that has large values on points from $P$ and small values on points from $Q$
    - Test statistic is the difference between the mean function values on the two samples
    - Large values of the test statistic provide evidence of $P \neq Q$
    
---

# Formal definition

$$MMD(\mathcal{F}, P, Q) = \sup_{f \in \mathcal{F}}\left( \mathbb{E}_Pf(X) - \mathbb{E}_Qf(Y)\right)$$

Empirical estimate:

$$\sup_{f\in \mathcal{F}}\left( \frac{1}{m}\sum_{i = 1}^mf(x_i) - \frac{1}{n}\sum_{i = 1}^n f(y_i)\right)$$

- $\mathcal{F}$ is the unit ball in a universal RKHS
    - universal means being dense in the space of bounded continuous functions




---

## Project Plan

- Verify results in `r Citep(bib, "gretton2012kernel")`  
- Supplement with reading from Berlinet & Thomas's "Reproducing kernel Hilbert spaces in probability and statistics" text   
- Examine software implementations of MMD calculations

*If time permits*

- Examine software applications to biological data
- Apply MMD to existing RNAseq data







---

## References

```{r refs, echo=FALSE, results="asis"}
PrintBibliography(bib)
```



---

class: center, middle

# Thanks!

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

